---
title: "Udacity Diamonds Exercise"
author: "Jonathan Morken"
date: "February 15, 2016"
output: html_document
---

The following is my full notes/results from performing the Udacity Diamonds Promblem set.

###Section One - Answers to introductory questions:
There are 53940 observations and 10 variables as can be seen with the str f(x).  3 of the variables are ordered factors. 
```{r,echo=FALSE}
#Load the Libraries
library(ggplot2)
data(diamonds)
library(moments)
library(ggplot2)
library(ggthemes)

#note, there are 53940 obs and 10 variables as can be seen with the str f(x).  All other answers to questions are here, as well.
str(diamonds)
```



###Section Two - Create a histogram of the price of all the diamonds in the diamond data set:
You will note the shape of the curve.  It is non-normal with a rather long tail.  Basic statistics appear below the graph:
```{r, echo=FALSE, figwidth = 12, warning=FALSE,fig.align='center'}
G1 <- ggplot(data = subset(diamonds,!is.na(price)), aes(x = price))

G1 +
  geom_histogram(binwidth = 500, colour = "white", fill = "darkgreen") +
  scale_x_continuous(limits = c(0,20000),breaks = seq(0,20000,5000)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")

```

**Note that the above receives a warning message that R Removed 2 rows containing missing values (geom_bar); however, I have looked for nulls and NAs throughout the dataset to no avail.  What am I missing?  

Summary Stats | #
------- | --------
Median | `r median(diamonds$price)`
Mean | `r round(mean(diamonds$price),digits = 2)` 
Skewness | `r round(skewness(diamonds$price),digits = 2)` 
Kurtosis |  `r round(kurtosis(diamonds$price),digits = 2)`


###Section Three - Answers to Diamond Count Questions:
How many diamonds cost less than $500: `r sum(diamonds$price < 500)`  
How many diamonds cost less than $250: `r sum(diamonds$price < 250)`  
How many diamonds cost $15,000 or more: `r sum(diamonds$price >= 15000)`


###Section Four - Exploring the Largest Peak in the Histogram:
Udacity Suggestions in the order they appear below: changing the limits (Dark Green), binwidths (Green), and breaks (Light Green).  
```{r, echo=FALSE, fig.width=3, warning=FALSE}

G1 +
  geom_histogram(binwidth = 250, colour = "white", fill = "darkgreen") +
  scale_x_continuous(limits = c(0,1500),breaks = seq(0,1500,500)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")

G1 +
  geom_histogram(binwidth = 50, colour = "black", fill = "green") +
  scale_x_continuous(limits = c(0,1500),breaks = seq(0,1500,500)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")

G1 +
  geom_histogram(binwidth = 50, colour = "black", fill = "lightgreen") +
  scale_x_continuous(limits = c(0,1500),breaks = seq(0,1500,150)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")
```

As apparent here, the most common price for diamonds is ~ $700.  

###Section Five - Break out the Histogram of Prices by Cut:
```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
G1+
  geom_histogram(binwidth = 500, colour = "white", fill = "darkgreen") +
  facet_grid(.~cut) +
  scale_x_continuous(limits = c(0,20000),breaks = seq(0,20000,5000)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")
```


###Section Six - Answer questions by Cut:
Max followed by Min and then Median
```{r, echo=FALSE}
by(diamonds$price,diamonds$cut,max)
by(diamonds$price,diamonds$cut,min)
by(diamonds$price,diamonds$cut,median)
```

###Section Seven - add a parameter to facet_wrap so that the y-axis in the histograms are not fixed.
```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
G1 +
  geom_histogram(binwidth = 500, colour = "white", fill = "darkgreen") +
  facet_wrap(~cut, scales="free_y") +
  scale_x_continuous(limits = c(0,20000),breaks = seq(0,20000,5000)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")
```

###Section Eight - Create a histogram of price per carat and facet wrap it by cut.
Additional Instructions from site - Adjust the bin width and transform the scale of the x-axis using log10.

```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
G1 +
  geom_histogram(binwidth = 500, colour = "white", fill = "darkgreen") +
  facet_grid(carat ~ cut, scales="free_y") +
  scale_x_continuous(limits = c(0,20000),breaks = seq(0,20000,5000)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")

```

