---
title: "Udacity Diamonds Exercise"
author: "Jonathan Morken"
date: "February 15, 2016"
output: html_document
---

The following is my set of notes/results from performing the first Udacity Diamonds problem set.

###Section One - Answers to introductory questions:
There are 53940 observations and 10 variables as can be seen with the str f(x).  3 of the variables are ordered factors. 
```{r,echo=FALSE,message = FALSE, warning=FALSE}
#Load the Libraries
library(ggplot2)
data(diamonds)
library(moments)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(plotly)

#note, there are 53940 obs and 10 variables as can be seen with the str f(x).  All other answers to questions are here, as well.
str(diamonds)
```



###Section Two - Create a histogram of the price of all the diamonds in the diamond data set:
You will note the shape of the curve.  It is non-normal with a rather long tail.  Basic statistics to answer section questions appear below the graph:
```{r, echo=FALSE, figwidth = 12, warning=FALSE,fig.align='center'}
G1 <- ggplot(data = subset(diamonds,!is.na(price)), aes(x = price))

G1 +
  geom_histogram(binwidth = 500, colour = "white", fill = "darkgreen") +
  scale_x_continuous(limits = c(0,20000),breaks = seq(0,20000,5000)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")

```

**Note that the above receives a warning message that R "Removed 2 rows containing missing values (geom_bar); however, I have looked for nulls and NAs throughout the data set to no avail.  What am I missing?  

Summary Stats | #
------- | --------
Median | `r median(diamonds$price)`
Mean | `r round(mean(diamonds$price),digits = 2)` 
Skewness | `r round(skewness(diamonds$price),digits = 2)` 
Kurtosis |  `r round(kurtosis(diamonds$price),digits = 2)`


###Section Three - Answers to Diamond Count Questions:
How many diamonds cost less than $500: `r sum(diamonds$price < 500)`  
How many diamonds cost less than $250: `r sum(diamonds$price < 250)`  
How many diamonds cost $15,000 or more: `r sum(diamonds$price >= 15000)`


###Section Four - Exploring the Largest Peak in the Histogram:
Udacity directions included [in order below]: changing the limits (Dark Green), bin-widths (Green), and breaks (Light Green).  
```{r, echo=FALSE, fig.width=3, warning=FALSE}

G1 +
  geom_histogram(binwidth = 250, colour = "white", fill = "darkgreen") +
  scale_x_continuous(limits = c(0,1500),breaks = seq(0,1500,500)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")

G1 +
  geom_histogram(binwidth = 50, colour = "black", fill = "green") +
  scale_x_continuous(limits = c(0,1500),breaks = seq(0,1500,500)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")

G1 +
  geom_histogram(binwidth = 50, colour = "black", fill = "lightgreen") +
  scale_x_continuous(limits = c(0,1500),breaks = seq(0,1500,150)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")
```

As apparent here, the most common price for a diamond is ~ $700, with a steep drop off below that.  

###Section Five - Break out the Histogram of Prices by Cut:
```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
G1+
  geom_histogram(binwidth = 500, colour = "white", fill = "darkgreen") +
  facet_grid(.~cut) +
  scale_x_continuous(limits = c(0,20000),breaks = seq(0,20000,5000)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")
```
No observations were requested, but it is interesting to me how uneven the distribution is between the various cuts.  There are only 1,610 fair and 4,906 good diamonds while their are 13,791 and 21,551 premium and ideal diamonds.  If this is meant to be representative of the normal population, De'Beers clearly does a good job selling it's "ideal."

###Section Six - Answer for max, min, and median questions by Cut, in that order:
```{r, echo=FALSE}
by(diamonds$price,diamonds$cut,max)
by(diamonds$price,diamonds$cut,min)
by(diamonds$price,diamonds$cut,median)
```

###Section Seven - Add a parameter to facet_wrap so that the y-axis in the histograms are not fixed.
```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
G1 +
  geom_histogram(binwidth = 500, colour = "white", fill = "darkgreen") +
  facet_wrap(~cut, scales="free_y") +
  scale_x_continuous(limits = c(0,20000),breaks = seq(0,20000,5000)) +
  xlab("Price in Dollars") +
  ylab("# of Diamonds")
```

I can't imagine using the above very often as it obscures relative size.

###Section Eight - Create a histogram of price per carat and facet wrap it by cut.
Additional Instructions from site - Adjust the bin width and transform the scale of the x-axis using log10.

```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
ggplot(data = subset(diamonds,!is.na(price)), aes(x = price/carat)) +
  geom_histogram(binwidth = .05, colour = "white", fill = "darkgreen") +
  facet_wrap(~ cut) +
  scale_x_log10() +
  xlab("Price per Carat in Dollars") +
  ylab("# of Diamonds")

```

I chose to remove the free scale y axis to see relative frequency of each price/carat more clearly.

### Section Nine - Investigate the price of diamonds using box plots.
Additional instructions - Also use numerical summaries and one of the following categorical variables: cut, clarity, or color. Then use a box plot to illustrate price/carat by color.
Note: I use the Plotly package for 2 base cases and the price/carat illustration here for experimentation and variety.  
  
####Price by cut:
```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
plot_ly(diamonds, y = price, color = cut, type = "box")
```
  
####Price by color:
```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
plot_ly(diamonds, y = price, color = color, type = "box")
```
  
####Price/carat by color:
```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
plot_ly(diamonds, y = price/carat, color = color, type = "box")
```

####Price by cut and color:
```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
ggplot(data = diamonds, aes(y = price, x = cut, fill = color)) +
  geom_boxplot() +
  theme_few()+
  scale_fill_pander() +
  ylab("Price per Diamond") +
  xlab("Cut of Diamond")

```


### Section 10 - Answers to section questions (IQR answers can be found by hovering over the appropriate boxplots above:
```{r,echo=FALSE}
iqrford <- subset(diamonds,color == 'D')
iqrforj <- subset(diamonds,color == 'J')
```

The 25th and 75th percentiles for colors D and J can be found by hovering over the box-plots above.  
IQR for color D: `r IQR(iqrford$price)`  
IQR for color J: `r IQR(iqrforj$price)`

### Section 11 - Investigate carat using a frequency polygon.
Additional directions - Use different bin widths to see how the polygon changes.
```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
g2 <- ggplot(data = diamonds, aes(x = carat)) +
  geom_freqpoly(binwidth = .5, colour = "darkgreen") +
  scale_x_continuous(limits = c(0,5),breaks = seq(0,5,.5)) +
  ylab("# of Diamonds") +
  xlab("Carat (i.e. Weight)")

g3 <- ggplot(data = diamonds, aes(x = carat)) +
  geom_freqpoly(binwidth = .1, colour = "darkgreen") +
  scale_x_continuous(limits = c(0,5),breaks = seq(0,5,.1)) +
  xlab("Carat (i.e. Weight)") +
  ylab("# of Diamonds")
```

The first graph shows the polygon with a bin-width of .5.  This obscures the variance in frequency within each carat.  The second graph shows the polygon with a bin-width of .1.  This allows the user to see how the market has pushed suppliers to produce more diamonds with "round" figures like 1/3 carat, 1/2 carat 3/4 carat, etc.
```{r,echo=FALSE, fig.width=10,warning=FALSE,fig.align='center'}
(ggplotly(g2))

(ggplotly(g3))
```
Answers to section questions:
What carat sizes have a frequency greater than 2000?  Only .3 and 1.01 from the list supplied.
