Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
library(plotly)
pf <- read.csv('pseudo_facebook.tsv',sep = '\t')

qplot(x = age, y = friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes: 

```{r ggplot Syntax}
g1 <- ggplot(data = pf, aes(x = age, y = friend_count))

g1 + geom_point(alpha = 1/20) +
  xlim(13,90)



summary(pf$age)
```

***

### Overplotting
Notes: Alpha represents transparency.  Jitter adds noise to distribution.

```{r Overplotting}
g1 + geom_jitter(alpha = 1/20) +
  xlim(13,90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans() Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!
Notes: Jitter doesn't work b/c of possible negative numbers with sqrt.  Must tell it to only "jitter" the age axis via the position syntax.

```{r Coord_trans()}
g1 + geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13,90) +
  coord_trans(y = "sqrt")
```

### Examine relationship between friendships_initiated (y) and age (x)

```{r}
g2 <- ggplot(data = pf,aes(x = age,y = friendships_initiated)) +
  geom_point()

g2 + geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13,90) +
  coord_trans(y = "sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)
age_groups <- group_by(pf,age)

pf.fc_by_age <- summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age,age)

head(pf.fc_by_age)


## Conditional Means alternate code:

pf.fc_by_age2 <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) %>%
  arrange(age)

head(pf.fc_by_age2,20)


```

Create your plot!

```{r Conditional Means Plot}
ggplot(data = pf.fc_by_age,aes(x = age, y = friend_count_mean))+
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes: Adding a line on top of existing plot.   

```{r Overlaying Summaries with Raw Data}
g1 + geom_point(alpha = 1/20,
                position = position_jitter(h = 0),
                color = "Orange") + 
  coord_cartesian(xlim = c(13,70),ylim = c(0,1000)) +
  geom_line(stat = 'Summary', fun.y = mean) +
  geom_line(stat = 'Summary', fun.y = quantile, fun.args = list(probs = .9), linetype = 2, color = 'blue') +
  geom_line(stat = 'Summary', fun.y = quantile, fun.args = list(probs = .1), linetype = 2, color = 'blue')
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age,pf$friend_count,method = "pearson")
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf,age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes: how to add a regression line to scatterplots.

```{r}
ggplot(data = pf,aes(y = likes_received, x = www_likes_received)) +
  geom_point()+
  xlim(0,quantile(pf$www_likes_received,0.95)) +
  ylim(0,quantile(pf$likes_received,0.95)) +
  geom_smooth(method = "lm", color = "red")
  
cor.test(pf$www_likes_received,pf$likes_received,method = "pearson")       
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}
data("Mitchell")
str(Mitchell)

ggplot(data = Mitchell,aes(x = Month, y = Temp)) +
  geom_point()+
  scale_x_discrete(breaks = seq(0,203,12))

cor.test(Mitchell$Month,Mitchell$Temp, method = "pearson")
range(Mitchell$Month)

## Method for graphing each year on top of each other:
ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+ 
  geom_point()

install.packages("energy")
library(energy)
x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(data = pf.fc_by_age,aes(x = age,y = friend_count_mean)) +
  geom_line()


```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- (pf$age + (1-(pf$dob_month/12)))
```

Programming Assignment
```{r Programming Assignment}
age_months_groups <- group_by(pf,age_with_months)

pf.fc_by_age_months <- summarise(age_months_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age,age_with_months)

head(pf.fc_by_age_months)



```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(data = subset(pf.fc_by_age_months,age_with_months < 70),aes(x = age_with_months,y = friend_count_mean)) +
  geom_line() 
  
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
library(gridExtra)
p1 <- ggplot(data = subset(pf.fc_by_age,age<71),aes(x = age,y = friend_count_mean)) +
  geom_line() +
  geom_smooth(color = "red")

p2 <- ggplot(data = subset(pf.fc_by_age_months,age_with_months < 70),aes(x = age_with_months,y = friend_count_mean)) +
  geom_line() +
  geom_smooth(color = "red")

grid.arrange(p2,p1,p3,ncol = 1)

p3 <- ggplot(data = subset(pf.fc_by_age,age<71),aes(x = round(age/5)*5,y = friend_count_mean))+
  geom_line(stat = "summary",fun.y=mean)

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

